---
title: "DATA 607 - Homework 9"
author: "Joshua Sturm"
date: "October 27, 2017"
output:
  html_document:
    code_folding: show
    toc: yes
    toc_float: yes
    smart: no
  pdf_document:
    keep_tex: yes
always_allow_html: yes
---

## Task

The New York Times web site provides a rich set of APIs, as described here:  http://developer.nytimes.com/docs.\
You'll need to start by signing up for an API key. Your task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and transform it to an R dataframe. 
 
## Load libraries
```{r loadlibraries, warning = F, message = F}
library(tidyverse)
library(jsonlite)
library(DT)
library(RCurl)
library(stringr)
```

## Methodology

I decided to use the Book API,located at https://developer.nytimes.com/books_api.json, which has data on the Times' best-sellers list.

```{r apikey, echo = F}
key <- "0b71fd02e63d4e6bb9f32fe6e519d63a"
```

## Example

```{r example}

# Base url:
base.url <- "https://api.nytimes.com/svc/books/v3/"
recent.best.url <- "lists/best-sellers/history.json?"
```

```{r exampleone2, echo = F}
recent.url <- paste0(base.url, recent.best.url, 'api-key=', key)
```

```{r exampleone3}
recent.get <- getURL(recent.url)
```

This is in a pretty ugly, unreadable format, so let's clean it up.

```{r exampleoneclean}
recent.clean <- fromJSON(recent.get) %>%
  as.data.frame() %>%
  unnest(results.isbns) %>%
  select(2, 4, 5, 7, 11:13)
names(recent.clean) <- str_replace_all(names(recent.clean), "results.", "")
datatable(recent.clean)
```

