---
title: "DATA 607 - Homework 7"
author: "Joshua Sturm"
date: "October 10, 2017"
output:
  html_document:
    code_folding: show
    toc: yes
    toc_float: yes
    smart: no
  pdf_document:
    keep_tex: yes
always_allow_html: yes
---
# Task

Pick three of your favorite books on one of your favorite subjects. At least one of the books should have more
than one author. For each book, include the title, authors, and two or three other attributes that you find
interesting.

Take the information that you've selected about these three books, and separately create three files which
store the book's information in HTML (using an html table), XML, and JSON formats (e.g. "books.html",
"books.xml", and "books.json"). To help you better understand the different file structures, I'd prefer that you
create each of these files "by hand" unless you're already very comfortable with the file formats.

Write R code, using your packages of choice, to load the information from each of the three sources into
separate R data frames. Are the three data frames identical?


## Approach

I began by picking three books on the subject of mathematics. For each book, I include the Title, Author, ISBN#, Publisher, date of publication, and number of pages. 

## Import the files from github

```{r libraries}
library(RCurl)
library(XML)
library(DT)
```

```{r getHtml}
# https://stackoverflow.com/questions/1395528/scraping-html-tables-into-r-data-frames-using-the-xml-package/1401367#1401367

bookHtml <- getURL("https://raw.githubusercontent.com/JoshuaSturm/CUNY_MSDS/master/Fall_2017/DATA_607/DATA_607_Homework_7/DATA_607_Homework_7_HTML.html?token=AVfc-72VfHxyfl89TM0FWLH_pYAS4mEYks5Z5j7MwA%3D%3D")
  
bookHtml <- htmlTreeParse(bookHtml, error=function(...){}, useInternalNodes = TRUE)


tablehead <- xpathSApply(bookHtml, "//table/tr/th", xmlValue)
results <- xpathSApply(bookHtml, "//table/tr/td", xmlValue)

bookHtml <- as.data.frame(matrix(results, ncol = 6, byrow = T))
names(bookHtml) <- tablehead

datatable(bookHtml)
```