---
title: "DATA607 - Assignment 1"
author: "Joshua Sturm"
date: "August 29, 2017"
output:
  html_document:
    code_folding: show
---


## DATA 607 - Assignment 1

Your task is to study the dataset and the associated description of the data (i.e. "data dictionary"). You may need
to look around a bit, but it's there! You should take the data, and create a data frame with a subset of the columns
in the dataset. You should include the column that indicates edible or poisonous and three or four other columns.
You should also add meaningful column names and replace the abbreviations used in the data-for example, in the
appropriate column, "e" might become "edible." Your deliverable is the R code to perform these transformation
tasks.

### Import the mushroom database

```{r import-data, eval=TRUE, echo=TRUE, stringsAsFactors=FALSE}
mushroomDB<- read.csv("https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data")
```

##### Narrow the database to include only the following characteristics:
##### `Class`, `Cap Color`, `Bruised`, `Odor`, `Gill Color`, `Stalk Color`, `Veil Color`, `Spore Color`, `Population`, and `Habitat`

```{r eval=TRUE}
mDB2 <- subset(mushroomDB, select = c(1,4,5,6,10,15,18,21,22,23))
```

#### Using the information from (~/agaricus-lepiota.names), rename the columns/variables 

```{r eval=TRUE}
colnames(mDB2) <- c("Class","Cap Color","Bruised","Odor","Gill Color","Stalk Color","Veil Color","Spore Color","Population","Habitat")
```

```{r}
mDB2[,1] <- ifelse(mDB2[,1] == "e", "Edible", ifelse(mDB2[,1] == "p", "Poisonous", 99))
mDB2[,2] <- ifelse(mDB2[,2] == "n", "Brown",
              ifelse(mDB2[,2] == "b", "Buff",
              ifelse(mDB2[,2] == "c", "Cinnamon",
              ifelse(mDB2[,2] == "g", "Gray",
              ifelse(mDB2[,2] == "r", "Green",
              ifelse(mDB2[,2] == "p", "Pink",
              ifelse(mDB2[,2] == "u", "Purple",
              ifelse(mDB2[,2] == "e", "Red",
              ifelse(mDB2[,2] == "w", "White",
              ifelse(mDB2[,2] == "y", "Yellow", "N/A")
                  )))))))))
mDB2[,3] <- ifelse(mDB2[,3] == "t", "Yes", ifelse(mDB2[,3] == "f", "No", "N/A"))
mDB2[,4] <- ifelse(mDB2[,4] == "a", "Almond",
              ifelse(mDB2[,4] == "l", "Anise",
              ifelse(mDB2[,4] == "c", "Creosote",
              ifelse(mDB2[,4] == "y", "Fishy",
              ifelse(mDB2[,4] == "f", "Foul",
              ifelse(mDB2[,4] == "m", "Musty",
              ifelse(mDB2[,4] == "n", "None",
              ifelse(mDB2[,4] == "p", "Pungent",
              ifelse(mDB2[,4] == "s", "Spicy", "N/A"
                  )))))))))
mDB2[,5] <- ifelse(mDB2[,5] == "k", "Black",
              ifelse(mDB2[,5] == "n", "Brown",
              ifelse(mDB2[,5] == "b", "Buff",
              ifelse(mDB2[,5] == "h", "Chocolate",
              ifelse(mDB2[,5] == "g", "Gray",
              ifelse(mDB2[,5] == "r", "Green",
              ifelse(mDB2[,5] == "o", "Orange",
              ifelse(mDB2[,5] == "p", "Pink",
              ifelse(mDB2[,5] == "u", "Purple",
              ifelse(mDB2[,5] == "e", "Red",
              ifelse(mDB2[,5] == "w", "White",
              ifelse(mDB2[,5] == "y", "Yellow", "N/A"
                  ))))))))))))
mDB2[,6] <- ifelse(mDB2[,6] == "n", "Brown",
              ifelse(mDB2[,6] == "b", "Buff",
              ifelse(mDB2[,6] == "c", "Cinnamon",
              ifelse(mDB2[,6] == "g", "Gray",
              ifelse(mDB2[,6] == "o", "Orange",
              ifelse(mDB2[,6] == "p", "Pink",
              ifelse(mDB2[,6] == "e", "Red",
              ifelse(mDB2[,6] == "w", "White",
              ifelse(mDB2[,6] == "y", "Yellow", "N/A"
                  )))))))))
mDB2[,7] <- ifelse(mDB2[,7] == "n", "Brown",
              ifelse(mDB2[,7] == "o", "Orange",
              ifelse(mDB2[,7] == "w", "White",
              ifelse(mDB2[,7] == "y", "Yellow", "N/A"
                  ))))
mDB2[,8] <- ifelse(mDB2[,8] == "k", "Black",
              ifelse(mDB2[,8] == "n", "Brown",
              ifelse(mDB2[,8] == "b", "Buff",
              ifelse(mDB2[,8] == "h", "Chocolate",
              ifelse(mDB2[,8] == "r", "Green",
              ifelse(mDB2[,8] == "o", "Orange",
              ifelse(mDB2[,8] == "u", "Purple",
              ifelse(mDB2[,8] == "w", "White",
              ifelse(mDB2[,8] == "y", "Yellow", "N/A"
                  )))))))))
mDB2[,9] <- ifelse(mDB2[,9] == "a", "Abundant",
              ifelse(mDB2[,9] == "c", "Clustered",
              ifelse(mDB2[,9] == "n", "Numerous",
              ifelse(mDB2[,9] == "s", "Scattered",
              ifelse(mDB2[,9] == "v", "Several",
              ifelse(mDB2[,9] == "y", "Solitary", "N/A"
                  ))))))
mDB2[,10] <- ifelse(mDB2[,10] == "g", "Grasses",
              ifelse(mDB2[,10] == "l", "Leaves",
              ifelse(mDB2[,10] == "m", "Meadows",
              ifelse(mDB2[,10] == "p", "Paths",
              ifelse(mDB2[,10] == "u", "Urban",
              ifelse(mDB2[,10] == "w", "Waste",
              ifelse(mDB2[,10] == "d", "Woods", "N/A"
                  )))))))
```




#### Use the DataTable package to simplify, tidy, and speed up the data presentation. 
#### I prefer it over rstudio's `df_print`

```{r}
 library(DT)
 datatable(mDB2, extensions = 'Scroller', options = list(
   deferRender = TRUE,
   scrollY = 200,
   scroller = TRUE
          ))
```