---
title: "DATA 607 - Homework 3"
author: "Joshua Sturm"
date: "09/14/2017"
output:
  html_document:
    code_folding: show
    toc: yes
    toc_float: yes
  pdf_document:
    keep_tex: yes
always_allow_html: yes
---
  
## Exercise 8.3

  **(a)** Use the tools of this chapter to rearrange the vector so that all elements conform to
the standard first_name last_name.

Solution:

```{r}
# Load the stringr library

library(stringr)

# Input the data

raw.data <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"   

# Extract just the names (by looking for letters)

name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
name

# Run a loop through the subset name. If there's a comma in that name, split it into a 1x2 vector, and switch the order. Lastly, 'trim' the whitespace from the left side.

for (i in 1:(length(name)))
{
  pat <- ","
  if (grepl(pat, name[i]) == 1)
  {
    separate_names <- unlist(str_split(name[i], ","))
    name[i] <- str_c(separate_names[2], separate_names[1], sep = " ")
  }
}
str_trim(name, side = "left")
```

### Exercise 8.3b
  **(b)** Construct a logical vector indicating whether a character has a title (i.e., Rev. and
Dr.).
  
Solution:

```{r}
# Search our subset for either of these titles

grep("(Rev.|Dr.)", name)

# Identify the two results

name[c(3,6)]
```

### Exercise 8.3c

  **(c)** Construct a logical vector indicating whether a character has a second name.
  
Solution:

```{r}
second_name <- str_count(fixed(name[c(1,2,4,5)]), "\\w+")
grep("3", second_name)
name[2]
```

## Exercise 8.4

  Describe the types of strings that conform to the following regular expressions and construct an example that is matched by the regular expression.
  
### 8.4a 

(a) [0-9]+ \\ \$\

```{r}
# Matches one or more digits followed by a dollar sign.
pat <- "[0-9]+\\$"
ex <- c("1$", "abc123$", "123$abc")
grepl(pat, ex)
```

### 8.4b

(b) \\ b[a-z]{1,4} \\ b\

```{r}
# Matches between 1 and 4 consecutive lowercase letters
pat <- "\\b[a-z]{1,4}\\b"
ex <- c("alpha", "beta", "gamma", "ZeTa", "eta")
grepl(pat, ex)
```

### 8.4c

(c) .*? \\ .txt\$\

```{r}
# Matches wildcard string ending in .txt
pat <- ".*?\\.txt$"
ex <- c("abc.txt", "123.txt", "123abc123.txt", "%#LMV#Ckdnf3i2.txt")
grepl(pat, ex)
```

### 8.4d

(d) \\ d{2}/ \\ d{2}/ \\ d{4}

```{r}
# Matches a string of digits in the format 12/34/5678
pat <- "\\d{2}/\\d{2}/\\d{4}"
ex <- c("98/76/5432", "1/2/3", "12/34/56")
grepl(pat, ex)
```

### 8.4e

(e) <(.+?)>.+?</ \\ 1>\

```{r}
# Matches 
```

## Exercise 8.9

The following code hides a secret message. Crack it with R and regular expressions.

clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0Tanwo
Uwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigO
d6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5
fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr

Solution:

My first guess was all the capital letters, since there were only a handful of them. Extracting each manually, I found this: CONGRATULATIONSYOUAREASUPERNERD. Now to do it in `R`:

```{r}
# Load the encrypted message 

encrypt <- "clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0Tanwo
Uwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigO
d6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5
fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr"

# Extract just the capital letters

decrypt <- unlist(str_extract_all(encrypt, "[[:upper:]]{1,}"))
decrypt
```